### 2023-08-13 tex/shared.tex

%
% Copyright (c) Meta Platforms, Inc. and affiliates.
%
% This source code is licensed under the MIT license found in the LICENSE file
% in the root directory of this source tree. 
%

% SIAM Shared Information Template
% This is information that is shared between the main document and any
% supplement. If no supplement is required, then this information can
% be included directly in the main document.

% Packages and macros go here
\usepackage{lipsum}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{epstopdf}
%%%%\usepackage{algorithmic}
\ifpdf
  \DeclareGraphicsExtensions{.eps,.pdf,.png,.jpg}
\else
  \DeclareGraphicsExtensions{.eps}
\fi

% Add a serial/Oxford comma by default.
\newcommand{\creflastconjunction}{, and~}

% Used for creating new theorem and remark environments
\newsiamremark{remark}{Remark}
\newsiamremark{hypothesis}{Hypothesis}
\crefname{hypothesis}{Hypothesis}{Hypotheses}
\newsiamthm{claim}{Claim}

% Sets running headers as well as PDF title and authors
\headers{An efficient algorithm for integer lattice reduction}
        {Fran\c{c}ois Charton, Kristin Lauter, Cathy Li, and Mark Tygert}

% Title. If the supplement option is on, then "Supplementary Material"
% is automatically inserted before the title.
\title{An efficient algorithm for integer lattice reduction\thanks{Submitted
to the editors DATE.
\funding{This work was funded by Meta Platforms, Inc.}}}

% Authors: full names plus addresses.
\author{Fran\c{c}ois Charton\thanks{Meta Platforms, Inc.,
  6 Rue M\'{e}nars 75002, Paris, France (\email{fcharton@meta.com}).}
\and Kristin Lauter\thanks{Meta Platforms, Inc.,
  1101 Dexter Ave N, Seattle, WA (\email{klauter@meta.com}).}
\and Cathy Li\thanks{Meta Platforms, Inc.,
  1101 Dexter Ave N, Seattle, WA (\email{cathyli@meta.com}).}
\and Mark Tygert\thanks{Meta Platforms, Inc.,
  1 Facebook Way, Menlo Park, CA (\email{tygert@meta.com}).}}

\usepackage{amsopn}
\DeclareMathOperator{\nint}{nint}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator{\bigoh}{\mathcal{O}}
\def\Id{\rm Id}

%%% Local Variables: 
%%% mode:latex
%%% TeX-master: "ex_article"
%%% End: